---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# datavis: R package for easy data visualization

<!-- badges: start -->

<!-- badges: end -->

The goal of datavis is to simplify the creation, annotation and export of bar plots and box plots of publication-grade quality. It is meant for academic researchers of any programming experience - newcomers and programming experts alike.

To a large extent, datavis wraps around the well-known [ggplot2 package](https://ggplot2.tidyverse.org/), but specializes in *very* specific tasks (only dodged bar and box plots), aiming to increase efficiency and accessibility to these tasks.

## Why datavis

Using the datavis functions, you can easily perform the following tasks with very few lines of code:

-   Create bar/box plots with summary statistics, optionally including individual data points.

-   Annotate with custom text above bars/boxes.

-   Customize plots aesthetically.

-   Export plots in various formats for any purpose.

## Installation

You can install datavis from [GitHub](https://github.com/) with:

```{r,eval=FALSE}
install.packages("devtools")
devtools::install_github("dimitriskokoretsis/datavis")
```

## Example

The following example demonstrates the use of all datavis functions to create, annotate and export a bar plot.

### Data import and plot creation

```{r}
# Data import using the fread function of the data.table package
demo.data.1 <- data.table::fread("guide/demo_data/demo_data_1.csv")

# Plot creation
library(datavis)

plot.1 <- demo.data.1 |>
  bar_point_plot(x="factor.1", # X axis grouping based on "factor.1" field
                 y="value", # Y axis value is "value" field
                 color.group="factor.2", # Color grouping based on "factor.2" field
                 x.axis="Factor 1", # Give a better title to x axis
                 y.axis="Value", # Give a better title to y axis
                 legend.title="Factor 2", # Give a better title to the legend
                 jitterwidth=0.7) # Adjust horizontal jitter of individual data points

plot.1
```

### Import of statistics data and plot annotation

```{r}
# Import of statistics data from Tukey's honest significant difference (HSD) test
demo.data.1.TukeyHSD <- data.table::fread("guide/demo_data/demo_data_1_TukeyHSD.csv")

demo.data.1.TukeyHSD
```

```{r}
# Annotation of original plot
plot.1.TukeyHSD <- plot.1 |>
  plot_stats(d=demo.data.1.TukeyHSD, # The data.frame containing the labels to be plotted.
             labels="HSDgroups", # The name of the labels column in the supplied data.frame.
             position="dodge") # Positioning of labels in the X dimension.
                               # See function's documentation for details.

plot.1.TukeyHSD
```

### Plot export

```{r}
plot.1.TukeyHSD |>
  plot_save(filepath="guide/demo_plots/plot_1_TukeyHSD", # Path to export files
            height=4,width=5)
```
